---
output:
  pdf_document: 
    keep_tex: yes
    latex_engine: xelatex
  word_document: default
  
header-includes:
   - \usepackage{xeCJK}
---

# Biostatistics Homework 4

By $\mathbb{L}$umi (张鹿鸣12112618)

\vspace{5mm}

## 1 Gene Expression Distribution

### 1.1)

```{r, fig.height=4, fig.width=6, fig.align='center'}
curve(dgamma(x, shape = 2, scale = 1) + dnorm(x, mean = 7, sd = 0.5) / 20,
      from = 0, to = 10,
      main = "Gene Expression Distribution",
      xlab = "Gene expression level", ylab = "Counts",
      xaxt = 'n', yaxt = 'n')
```

I think the gene expression pattern will follow a Gamma distribution, which in other words is a distribution skewed to the right. Because many genes are not significantly expressed, thus most of the genes will have a rather low expression level. But some genes are expressed in a very significant amount, because they might be essential for cell survival, or are the signatures for this specific cell type. So a few counts will lie in the high expression area. Thus I think a Gamma distribution with a small bump on the right will fit best.

### 1.2)

```{r, fig.width=6, fig.height=4, fig.align='center'}
curve(dnorm(x, mean = 0, sd = 1), 
      from = -4, to = 4,
      main = "Normal Distribution",
      xlab = "Gene expression level", ylab = "Cell counts",
      xaxt = 'n', yaxt = 'n')
```

I think the *Nanog* gene expression across the embryonic stem cell population should follow a normal distribution. Because they are all from the same embryonic stem cell population, their expression of a particular gene should not vary from a certain value vary dramatically, although there might be fluctuations.

### 1.3)

I think the distributions of 1.1) and 1.2) are not the same. 1.1) is the distribution of expression levels of all the genes in a single cell, it shows that genes are expressed differently in this cell. 1.2) is the distribution of the expression of a gene in multiple cells, it shows the general level of expression of this gene in all the cells.

## 2. Weaver ants

### 2.1)

```{r, fig.width=6, fig.height=4, fig.align='center'}
curve(dnorm(x, mean = 5, sd = 1.7) + dnorm(x, mean = 10, sd = 1.6),
      from = 0, to = 14,
      main = 'Population Distribution of Weaver Ants Body Length',
      xlab = 'Weaver Ant Body Length (mm)',
      yaxt = 'n', ylab = 'Counts')
```

Because there are two types of weaver ants with different body lengths, I suspect that the curve can have two peaks accounting for the two different types of ants. Among the two different types of ants, the body length should follow a normal distribution respectively, so the population distribution of the body length should simply be the addition of the two normal distributions.

### 2.2)

Since the sample is drawn randomly from the population, we can assume that the sample follows the same type of distribution of the population.

```{r, fig.width=6, fig.height=4, fig.align='center'}
hist(rnorm(312, mean = 5, sd = 1.7) + rnorm(313, mean = 10, sd = 1.6),
     main = 'Sample Distribution of Weaver Ants Body Length',
     xlab = 'Weaver Ant Body Length (mm)', xaxt = 'n',
     yaxt = 'n', ylab = 'Counts',
     breaks = 50)
```

### 2.3)

When you randomly choose multiple groups of 625 weaver ants, the mean body length should follow a normal distribution with $\mu = 7, \sigma^2 = (\frac{10}{25})^2$. Suppose $X_i$ is the random variable indicating the mean body length for the 625 weaver ants.

$$
P(X_i \leq 7.5) = P(Z \leq \frac{7.5 - \mu}{\sigma}) = P(Z \leq 1.25) \approx 0.8944
$$

So the probability of getting a mean from 625 weaver ants that is smaller than 7.5mm is 0.8944.

## 3 A Simple Random Sample

C

## 4 The Concept of Population

[ tick ] It is better to think of our population as an abstract concept.

[ tick ] The population is effectively infinite.

## 5 A Practical Simple Random Sampling Strategy

```{r}
sample(1:1000, 20)
```

But actually, I think just taking the first 20 participants is OK, if your 1000 participants are labeled randomly.

## 6 Triceps skinfold thickness

### 6.1)

Similar to problem 2 about the ants, suppose the mean triceps skinfold thickness of 25 people follows is the random variable $X$. Than $X$ should follow a normal distribution with parameters $\mu = 1.35, \sigma^2 = 0.01$.

$$
P(X \leq 1.05) = P(Z \leq \frac{1.05 - \mu}{0.1} )  = P(Z \leq -3) \approx 0.0013
$$

### 6.2)

From 6.1) we can see that, getting 25 random people and observing a mean value smaller than 1.05 is very rare ( $p < 0.0013$ ). From the possibility, you can say that it's a once in a thousand opportunity to get a random sample like this. We can also calculate the possibility to observe a mean larger than \$X = 0.92\$, which is $p = 0.99999$. It is highly unlikely to get a value smaller than 0.92cm. Thus we can say that these people who suffer from COPD are significantly different from the normal population.

## 7 Decrease Manufacturing Variability

### 7.1)

Suppose that the weights of bags of potato chips follows the random variable $X$.

$$
P(X \leq 330) = P(Z \leq \frac{330 - 362}{20} ) = P( Z\leq -1.6) \approx 0.0548
$$

### 7.2)

Suppose the random variable $Y$ describes the mean of the sample with size 100. $Y$ follows a normal distribution $Y \sim N(362, 4)$.

$$
P(360 < Y < 363) = P( -1 < Z < 0.5) = 0.5328
$$

### 7.3)

Because only 1% of bags weigh less than 330g, the Z score for 1% is $-2.326$.

$$
\begin{aligned}
\frac{330 - 362}{\sigma} &= -2.326 \\
\sigma &\approx 13.75
\end{aligned}
$$

## 8 A Hypothetical Linear PDF

### 8.1)

$$
\begin{aligned}
\int_0^1 (1 - \theta) + 2\theta x dx = [(1-\theta)x + \theta x^2 ]_0^1 = 1
\end{aligned}
$$

This shows that the area under the PDF is 1.

$$
\begin{aligned}
0 < &x<1 \\
1- \theta < &f_\mathbf{X}(x) < 1 + \theta
\end{aligned}
$$

Because $-1 < \theta < 1$, $0 < f_\mathbf{X}(x) < 2$.

So this probability density functions fulfills the basic properties of the PDF. That is, the area under the curve is 1, and all the density values are larger than 0.

### 8.2)

$$
F_X(x) = \int_0^x(1-\theta) + 2\theta t dt = (1-\theta)x + \theta x^2
$$

So

$$
F_X(x) = \begin{cases}
1 &, x > 1 \\
(1-\theta) x + \theta x^2 &, 1 > x > 0 \\
0 &, x\leq0
\end{cases}
$$

### 8.3)

$$
\begin{aligned}
E(X) &= \int_0^1 x(1-\theta) + 2\theta x^2 dx \\
&= \frac{(1-\theta)x^2}{2} + \frac{2\theta x^3}{3} \rbrack_0^1 \\
& = \frac{3 + \theta}{6}
\end{aligned}
$$

### 8.4)

$$
\mathcal L(\theta; x_1,x_2, ..., x_n) = \prod_{i=1}^n [(1-\theta) + 2\theta x_i]
$$

So, $\ell(\theta;x) = \ln{\mathcal L(\theta; x)}$

$$
\ell (\theta; x) = \sum_{i=1}^n\ln{[(1-\theta) + 2\theta x_i]}
$$

### 8.5)

$$
\begin{aligned}
\frac{d \ell}{d \theta} = \sum_{i=1}^n \frac{2x_i - 1}{(1-\theta) + 2\theta x_i} = 0 \\
\hat{\theta} = t \\
\sum_{i=1}^n \frac{2x_i - 1}{(1-\hat \theta) + 2\hat \theta x_i}= \sum_{i=1}^n \frac{2x_i - 1}{(1-t) + 2t x_i} = 0
\end{aligned}
$$

So we choose B.

## 9 Waiting Time Between Text Messages

$$
\begin{aligned}
\mathcal{L}(\lambda; x) &= \prod_{i=1}^n\lambda e^{-\lambda x_i} = \lambda^n e^{-\lambda \sum_{i=1}^nx_i} \\
\ell (\lambda;x) &= \ln{\mathcal{L}(\lambda;x)} \\
&= n \ln \lambda - \lambda \sum_{i=1}^n x_i \\
\frac{d \ell}{d \lambda} &= 0 = \frac{n}{\hat \lambda}- \sum_{i=1}^n x_i \\
\hat \lambda &= \frac{n}{\sum_{i=1}^n x_i}
\end{aligned}
$$

So the maximum likelihood estimation of $\lambda$ is $\frac{n}{\sum_{i=1}^n x_i}$.
