---
output:
  pdf_document: 
    keep_tex: yes
    latex_engine: xelatex
  word_document: default
  
header-includes:
   - \usepackage{xeCJK}
---

# Biostatistics Homework 5

By $\mathbb{L}$umi (张鹿鸣12112618)

\vspace{5mm}

## 1. One-sided confidence interval 1

In this question, we are interested in one sided confidence intervals. Thus we want to find:

$$
\text P (\mu \leq b) = 0.95
$$

Since the normal distribution is symmetric, and our population luckily follows a normal distribution with a know standard deviation, the one sided confidence interval for the lower tail and the upper bond will be just symmetric around $\bar X$. The margin of error can be calculated with a z score.

```{r}
qnorm(0.95)
```

We get from R that when $z = 1.645$, $\text P (Z < z) = 0.95$. We also know that:

$$
Z = \frac{\bar X - \mu}{\sigma / \sqrt{n}}
$$

Thus the one sided confidence interval of the lower tail will be $\text{C.I.} = (- \infty, \bar X + 1.645 \frac{\sigma}{\sqrt{n}}]$.

Similarly, the one sided confidence interval of the upper tail will be $\text{C.I.} = [\bar X - 1.645 \frac{\sigma}{\sqrt{n}}, + \infty)$.

## 2. One-sided confidence interval 2

Different from question 1, although our population still follows a normal distribution, this time, we don't know anything about it's variance. Thus, we need to substitute the population variance with the sample variance. Our random variable $\frac{\bar X - \mu}{S/ \sqrt{n}}$ follows a T-distribution. We want to find:

$$
\text{P}(\mu \leq b) = 0.95
$$

The T-distribution is also symmetric, so the upper tail and the lower tail is symmetric around $\bar X$. The margin of error can be calculated from the t score.

If we want to calculate the confidence interval for the lower tail, the t score is $t_{0.05, v - 1}$. We also know that:

$$
t = \frac{\bar X - \mu}{s / \sqrt{n}}
$$

Thus the confidence interval for the lower tail will be $\text{C.I.} = (- \infty, \bar X + t_{0.05,v-1} \frac{s}{\sqrt{n}}]$.

Similarly, the confidence interval for the upper tail will be $\text{C.I.} =[\bar X + t_{0.95,v-1} \frac{s}{\sqrt{n}}, + \infty)=[\bar X - t_{0.05,v-1} \frac{s}{\sqrt{n}}, + \infty)$.

## 3.

T

F

$[0.7421, 0.8229]$ (See 3 supplement)

F

T

#### 3 supplement

We know from the problem that this is a test for proportion. Thus the confidence interval is:

```{r}
qnorm(0.975)
```

```{r}
313/400
```

$$
\text{C.I.} = [0.7825 - 1.96 \sqrt{\frac{(1-0.7825) \times 0.7825}{400}},0.7825 + 1.96 \sqrt{\frac{(1-0.7825) \times 0.7825}{400}}]
$$

Which is $[0.7421, 0.8229]$.

## 4 Null and Alternative Hypothesis

### 4.1)

$H_0$ : The rate of female student hadn't increased. $\pi \leq 7 \%$

$H_1$ : The rate of female student increased. $\pi > 7 \%$

### 4.2)

$H_0$ : The new technique is not significantly better or worse. $\pi = 75 \%$

$H_1$ : The new technique is significantly better or worse. $\pi \not = 75 \%$

### 4.3)

$H_0$ : The machine is not dispensing too much in medium drinks. $\mu \leq 530 \text{ml}$

$H_1$ : The machine is dispensing too much in medium drinks. $\mu > 530 \text{ml}$

### 4.4)

$H_0$ : The proportion of people who received the flu shot had not changed. $\pi = 48 \%$

$H_1$ : The proportion of people who received the flu shot had changed. $\pi \not = 48 \%$

## 5 The $p$-value

F

F

T

F

F

## 6 Blue M&M's candies

### 6.1)

D

### 6.2)

$H_0$: The plain M&M's candies do contain 24% blue ones. $\pi = 24 \%$

$H_1$: The plain M&M's candies does not contain 24% blue ones. $\pi \not= 24 \%$

### 6.3)

The test statistic, which should be a z-statistic, because we are testing on a proportion, and $np > 10, nq > 10$.

$$
z = \frac{p - \pi}{\sqrt{\frac{( 1 - \pi) \pi}{n}}} = \frac{18.4\% - 24\%}{\sqrt{\frac{(1-24\% ) \times 24\%}{500}}} \approx -2.932
$$

### 6.4)

The population follows a binomial distribution. Our sample satisfies $np > 10$ and $n(1-p) > 10$. Thus the test statistic approximately follows a standard normal distribution.

### 6.5)

There might be multiple reasons. I list some of my explanations below:

-   Li Lei is very lucky and got weird M&M's candies.

-   Li Lei bought the 500 candies from the same production batch, and they are not of good quality.

-   The M&M might change their policy on the proportion of blue candies.

-   Li Lei might have counting mistakes (e.g. He is colour blind)

### 6.6)

The margin of error is:

$$
z_{0.025} \sqrt{\frac{(1-p)p}{n}} = 1.96 \times \sqrt{\frac{(1-0.184) \times 0.184}{500}} = 0.03
$$

So the two sided 95% confidence interval is:

$$
\text{C.I.} = [15.4\%, 21.4 \%]
$$

### 6.7

```{r}
pnorm(-2.932) * 2
```

We can see that the $p$-value is 0.003 \< 0.01. Thus, Li Lei should reject the null hypothesis.

## 7 Renal Disease

### 7.1)

C

### 7.2)

$H_0$: The mean serum-creatinine level in the people who have taken the antibiotic is not significantly different from the normal people. $\mu = 1.0 \text{mg/dL}$

$H_1$: The mean serum-creatinine level in the people who have taken the antibiotic is significantly different from the normal people. $\mu \not= 1.0 \text{mg/dL}$

### 7.3)

Since we are comparing between the sample and the whole normal population, and the antibiotic does not change the dispersion of the serum-creatinine level, we can assume that a z statistic should be calculated.

$$
z = \frac{\bar x - \mu}{\sigma/ \sqrt{n}} = \frac{1.2 - 1.0}{0.4 / \sqrt{12}} = \sqrt{3} = 1.732
$$

### 7.4)

If $H_0$ was true, as mentioned above, the test statistics will follow a standard normal distribution.

### 7.5)

The margin of error is:

$$
\varepsilon = z_{0.025} \times \frac{\sigma}{\sqrt{n}} = 1.96 \times \frac{0.4}{\sqrt{12}} \approx 0.2263
$$

So the confidence interval is $\text{C.I.} = [0.9736, 1.4263]$.

### 7.6)

```{r}
pnorm(1.732, lower.tail = F) * 2
```

The \$p\$-value is 0.08 which is not smaller than 0.05, I think I will not reject $H_0$.

## 8 Serum Cholestrol Level

### 8.1)

From lecture 26, we now that the standard deviation in the sample is 46mg/100ml. We also know that we want a significance level of $\alpha = 0.05$. We can calculate the rejection threshold:

$$
x_{\text{reject}} = \bar x + z_{0.05} \times \frac{\sigma}{\sqrt{n}} = 180 + 1.645 \times \frac{46}{\sqrt{50}} \approx 190.7
$$

So the rejection threshold is 190.7mg/100ml.

The power, $1-\beta$, can be calculated from $\beta$, which is the possibility of not rejecting the null hypothesis when the null hypothesis is not true. First we calculate the test statistics for the population:

$$
z = \frac{x_{\text{reject}} - \mu}{\sigma / \sqrt{n}} = \frac{190.7 - 200}{46/\sqrt{50}} \approx -1.43
$$

Then we can calculate the probability of $P( \text{not reject } H_0 | H_0 \text{ is false})$ which is $\beta$.

```{r}
pnorm(-1.43)
```

So $\beta = 0.076$. The power is $1-\beta = 0.924$.

### 8.2)

Because I don't want a step by step calculation, so I'm just going to derive the general formula.

The test statistic we calculate are $z_\alpha$ and $z_{1-\beta} = -z_\beta$. From the previous example, we can see that:

$$
-z_\beta = \frac{\bar x + z_\alpha \times \frac{\sigma}{\sqrt{n}} - \mu}{\sigma / \sqrt{n}}
$$

With a bit of fumbling around, we can get the equation:

$$
n = (\frac{(z_\alpha + z_\beta) \sigma}{\mu - \bar x})^2
$$

So in our case, $\lceil n \rceil = 101$, so the required sample size is 101.

### 8.3)

B

## 9 Type I & II errors

### 9.1)

B

### 9.2)

C

### 9.3)

B

### 9.4)

D (But I think C is also ok, because we don't want to waste a lot of money to clean the pool.)
